---
title: "Analyzing Issue Scales"
# author: "Yen-Chieh Liao"
# date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyzing Issue Scales}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## 2.1.2 Example 1: 2009 European Election Study (French Module)

```{r}
library(asmcjr)
data(franceEES2009)
head(franceEES2009)
```

```{r}
library(basicspace)
result.france <- aldmck(franceEES2009, respondent=1, polarity=2,
     missing=c(77,88,89), verbose=FALSE)
```

```{r warning=TRUE, fig.cap="FIGURE 2.1: Aldrich-McKelvey Scaling of Left-Right Self- Placements of French Respondents (2009 European Election Study)"}
# plot stimuli locations in addition to ideal point density
library(ggplot2)
# plot density of ideal points
 plot_resphist(result.france, xlab="Left-Right")
```

```{r warning=TRUE, fig.cap="FIGURE 2.2: Aldrich-McKelvey Scaling of Left-Right Placements of French Political Parties (2009 European Election Study)"}
# plot stimuli locations in addition to ideal point density
plot_resphist(result.france, addStim=TRUE, xlab = "Left-Right") +
  theme(legend.position="bottom", aspect.ratio=1) +
  guides(shape = guide_legend(override.aes = list(size = 4), nrow=3)) +
 labs(shape="Party", colour="Party")
```

```{r warning=TRUE, fig.cap="FIGURE 2.3: Aldrich-McKelvey Scaling of Left-Right Placements of French Political Parties: Positive Weights (2009 European Election Study)"}
# isolate positive weights
plot_resphist(result.france, addStim=TRUE, weights="positive",
              xlab = "Left-Right")  +
  theme(legend.position="bottom", aspect.ratio=1) +
  guides(shape = guide_legend(override.aes = list(size = 4), nrow=3)) +
  labs(shape="Party", colour="Party")
```

```{r warning=TRUE, fig.cap="FIGURE 2.3: Aldrich-McKelvey Scaling of Left-Right Placements of French Political Parties: Negative Weights (2009 European Election Study)"}
# isolate positive weights
plot_resphist(result.france, addStim=TRUE, 
              weights="negative", xlab = "Left-Right")  +
  theme(legend.position="bottom", aspect.ratio=1) +
  guides(shape = guide_legend(override.aes = list(size = 4), nrow=3)) +
  labs(shape="Party", colour="Party")
```


## 2.1.3 Example 2: 1968 American National Election Study Urban Unrest and Vietnam War Scales

#### Urban Unrest
Running Bayesian Aldrich-Mckelvey
```{r message=FALSE, warning=FALSE}
# Loading 'nes1968_urbanunrest' 
data(nes1968_urbanunrest)
# Creating object with US president left-right dimensions
urban <- as.matrix(nes1968_urbanunrest[,-1])
# Running Bayesian Aldrich-Mckelvey scaling on President positions
library(basicspace)
result.urb <- aldmck(urban, polarity=2, respondent=5,
                     missing=c(8,9), verbose=FALSE)
summary(result.urb)
```
Extracting `vote.choice` Column
```{r message=FALSE, warning=FALSE}
# recode so that only Humphrey, Nixon and Wallace are present
vote <- car:::recode(nes1968_urbanunrest[,1], "3='Humphrey'; 5 = 'Nixon'; 6 = 'Wallace'; else=NA",
                     as.factor=FALSE)
# Convert vote to factor with appropriate levels
vote <- factor(vote, levels=c("Humphrey", "Nixon", "Wallace"))
```


```{r warning=TRUE, fig.cap="FIGURE 2.5: Aldrich-McKelvey Scaling of Urban Unrest Scale: Candidates and Voters (1968 American National Election Study)"}
# Plot population distribution by vote choice
plot_resphist(result.urb, groupVar=vote, addStim=TRUE,
              xlab="Liberal-Conservative")  +
  theme(legend.position="bottom", aspect.ratio=1)  +
  guides(shape = guide_legend(override.aes =
                                list(size = 4, color=c("gray25", "gray50", "gray75"))),
  colour = "none") +
  xlim(c(-2,2)) +
  labs(shape="Candidate")
```
#### Vietnam War Scales
```{r}
data(nes1968_vietnam)
vietnam <- as.matrix(nes1968_vietnam[,-1])
# Aldrich-Mckelvey function for vietnam dataset
result.viet <- aldmck(vietnam, polarity=2, respondent=5,
                      missing=c(8,9), verbose=FALSE)
summary(result.viet)
```

```{r warning=TRUE, fig.cap="FIGURE 2.6: Aldrich-McKelvey Scaling of Vietnam War Scale: Candidates and Voters (1968 American National Election Study)"}
# Plot population distribution by vote choice
plot_resphist(result.urb, groupVar=vote, addStim=TRUE,
              xlab="Liberal-Conservative")  +
  theme(legend.position="bottom", aspect.ratio=1)  +
  guides(shape = guide_legend(override.aes =
                                list(size = 4, color=c("gray25", "gray50", "gray75"))),
  colour = "none") +
  xlim(c(-2,2)) +
  labs(shape="Candidate")
```

```{r}
boot.france  <- boot.aldmck(franceEES2009,
                            polarity=2, respondent=1, missing=c(77,88,89),
                            verbose=FALSE, boot.args = list(R=100))
```


```{r warning=TRUE, fig.cap="FIGURE 2.7: Aldrich-McKelvey Scaling of Left-Right Placements of French Political Parties (2009 European Election Study) with Bootstrapped Standard Errors"}
library(ggplot2)
ggplot(boot.france$sumstats, aes(x = idealpt, y = stimulus)) +
  geom_point() +
  # geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.2) +
  xlab("Left-Right") +
  ylab(NULL) +
  theme_minimal() +
  theme(legend.position = "bottom", aspect.ratio = 1) 
```


## 2.2.1 Example 1: 2000 Convention Delegate Study

```{r}
library(basicspace)
data(CDS2000)
head(CDS2000[,5:8])
```
Blackbox syntax of Republican-Democrat left-right scale
```{r}
issues <- as.matrix(CDS2000[,5:14])
result.repdem <- blackbox(issues,
                          missing=99, dims=3, minscale=5, verbose=TRUE)
```
Recode party: Democrats = 1; Republicans = 2
```{r}
party <- car:::recode(CDS2000[,1],
                      "1='Democrat'; 2='Republican'; else=NA",
                      as.factor=TRUE)
```


```{r message=FALSE, warning=FALSE, fig.cap="FIGURE 2.8: Basic Space (Blackbox) Scaling of US Party Conven- tion Delegates (2000 Convention Delegate Study)"}
plot_blackbox(result.repdem, dims=c(1,2), groupVar=party,
              xlab= "First Dimension\n(Left-Right)",
              ylab="Second Dimension") +
  theme(legend.position="bottom", aspect.ratio=1) +
  guides(shape=guide_legend(override.aes=list(size=4))) +
  labs(colour="Party")
```

## 2.2.2 Example 2: 2010 Swedish Parliamentary Candidate Survey

```{r}
library(basicspace)
data(Sweden2010)
head(Sweden2010[,1:8])
```
Blacbox Scaling for Sweden issue scale
```{r}
# Extract issues scales and convert to numeric
issues.sweden <- as.matrix(Sweden2010[,7:56])
mode(issues.sweden) <- "numeric"
# Blacbox syntax for Sweden issue scale
result.sweden <- blackbox(issues.sweden, missing=8,
                          dims=3, minscale=5, verbose=FALSE)
```

```{r}
 # change polarity of scores
 if(result.sweden$individuals[[1]][13,1] < 0)
 result.sweden$individuals[[1]][,1] <-
     result.sweden$individuals[[1]][,1] * -1
 result.sweden$fits
```

```{r}
result.sweden$stimuli[[1]][16:25,]
```

```{r message=FALSE, warning=FALSE, fig.cap="FIGURE 2.11: Basic Space (Blackbox) Scaling of 2010 Swedish Parliamentary Candidate Data (Candidates by Party)"}
elected <- as.numeric(Sweden2010[,2])
party.name.sweden <- as.factor(Sweden2010[,3])
plot_resphist(result.sweden, groupVar=party.name.sweden, dim=1,
              scaleDensity=FALSE) +
   facet_wrap(~stimulus, ncol=2) +
   theme(legend.position="none") +
   scale_color_manual(values=rep("black", 10))
```

Density plot syntax and comparison of defeated/elected candidates

```{r message=FALSE, warning=FALSE, fig.cap="FIGURE 2.12: Basic Space (Blackbox) Scaling of 2010 Swedish Parliamentary Candidate Data (Elected and Defeated Candidates by Party)"}
# Keep only the parties of elected candidates, set others to NA
party.name.sweden[which(elected == 0)] <- NA
plot_resphist(result.sweden, groupVar=party.name.sweden, dim=1,  scaleDensity=FALSE) +
  facet_wrap(~stimulus, ncol=2) +
  theme(legend.position="none") +
  scale_color_manual(values=rep("black", 10))
```

## 2.2.3 Estimating Bootstrapped Standard Errors for Black Box Scaling

#### The 2010 Swedish parliamentary candidate data

```{r eval=TRUE, include=FALSE}
data(outbb)
```

```{r eval=FALSE, include=TRUE}
# Candidate point estimates blackbox scaling
outbb <- boot.blackbox(issues.sweden, missing=8, dims=3, minscale=5,
                      verbose=FALSE, posStimulus=13)
```

Matrix creation for Swedish candidates
```{r}
first.dim <- data.frame(
   point = result.sweden$individuals[[3]][,1],
   se = apply(outbb[,1,], 1, sd)
 )
first.dim$lower <- with(first.dim, point - 1.96*se)
first.dim$upper <- with(first.dim, point + 1.96*se)
first.dim$elected <- factor(elected, levels=c(0,1),
                            labels=c("Not Elected", "Elected"))
head(first.dim)
```

Plot for the distribution of first dimensions bootstrapped SE

```{r message=FALSE, warning=FALSE, fig.cap="FIGURE 2.13: Basic Space (Blackbox) Scaling of 2010 Swedish Parliamentary Candidate Data with Boostrapped Standard Errors (Elected and Defeated Candidates)"}
ggplot(first.dim, aes(x=se, group=elected)) +
  stat_density(geom="line", bw=.005) +
  facet_wrap(~elected) +
  theme(aspect.ratio=1) +
  xlab("Standard Error") +
  xlim(c(0,.2)) +
  theme_bw()
```

To determine whether the difference between the two distributions is statistically significant, we compute a permutation test for the difference in standard errors variances. 

```{r}
# Variance test syntax
library(perm)
levels(first.dim$elected) <- c("No", "Yes")
permTS(se ~ elected, data=first.dim,
      alternative="greater", method="exact.mc",
      control=permControl(nmc=10^4-1))

```

## 2.3.1 Example 1: 2000 and 2006 Comparative Study of Elec- toral Systems (Mexican Modules)

```{r}
data(mexicoCSES2000)
data(mexicoCSES2006)
head(mexicoCSES2000)
```
Blackbox syntax for two datasets, with data cleaning arguments

```{r eval=TRUE, include=FALSE}
data(result_2000)
data(result_2006)
```

```{r eval=FALSE, include=TRUE}
library(basicspace)
result_2000 <- blackbox_transpose(mexicoCSES2000, missing=99,
                                  dims=3, minscale=5, verbose=TRUE)
result_2006 <- blackbox_transpose(mexicoCSES2006, missing=99,
                                  dims=3, minscale=5, verbose=TRUE)
```
Multiplying here to avoid negative scores
```{r}
# Extract and transform dimensions for the year 2000
first.dim.2000 <- -1 * result_2000$stimuli[[2]][, 2]
second.dim.2000 <- result_2000$stimuli[[2]][, 3]

# Extract and transform dimensions for the year 2006
first.dim.2006 <- -1 * result_2006$stimuli[[2]][, 2]
second.dim.2006 <- result_2006$stimuli[[2]][, 3]

# Create a data frame for plotting
plot.df <- data.frame(
  dim1 = c(first.dim.2000, first.dim.2006),
  dim2 = c(second.dim.2000, second.dim.2006),
  year = rep(c(2000, 2006), c(length(first.dim.2000), length(first.dim.2006))),
  party = factor(c(rownames(result_2000$stimuli[[2]]), rownames(result_2006$stimuli[[2]])))
)

# Add nudge values for adjusting labels in the plot
plot.df$nudge_x <- c(0, 0, 0, 0, 0, -0.125, 0, 0, 0, 0.13, 0, 0, -0.225, 0)
plot.df$nudge_y <- c(-0.05, -0.05, 0.05, -0.05, -0.05, 0, -0.05, -0.05, -0.05, 0.03, 0.05, -0.05, -0.025, 0.05)

# Display the first few rows of the data frame
head(plot.df)

```

```{r message=FALSE, warning=FALSE, fig.cap="FIGURE 2.14: Basic Space (Blackbox Transpose) Scaling of Left- Right Placements of Mexican Political Parties (2000 and 2006 Com- parative Study of Electoral Systems)"}
ggplot(plot.df, aes(x=dim1, y=dim2, group=year)) +
  geom_point() +
  geom_text(aes(label=party), nudge_y=plot.df$nudge_y, size=3,
           nudge_x=plot.df$nudge_x, group=plot.df$year) +
  facet_wrap(~year) +
  xlim(-.55,1) +
  ylim(-.55,1) +
  theme_bw() +
  labs(x="First Dimension", y="Second Dimension")
```

## 2.3.2 Estimating Bootstrapped Standard Errors for Black Box Transpose Scaling
```{r}
rankings <- as.matrix(franceEES2009[,2:9])
mode(rankings) <- "numeric"
original <- blackbox_transpose(rankings,
                     missing=c(77,88,89), dims=3, minscale=5, verbose=FALSE)
```

```{r}
# Reverse check for the first stimulus
if (original$stimuli[[1]][1,2] > 0) {
    original$stimuli[[1]][,2] <- -1 * original$stimuli[[1]][,2]
}

# Print the fits from the original object
print(original$fits)
```

```{r eval=TRUE, include=FALSE}
data(outbbt)
```

```{r eval=FALSE, include=TRUE}
outbbt <- boot.blackbox_transpose(rankings, missing=c(77,88,89),
                                  dims=3, minscale=5, verbose=FALSE, R=5)
```

```{r}
# Create a data frame for the bootstrapped results
france.boot.bbt <- data.frame(
  party = colnames(rankings),
  point = original$stimuli[[1]][, 2],
  se = apply(outbbt[, 1, ], 1, sd, na.rm = TRUE)
)

# Calculate the confidence intervals
france.boot.bbt$lower <- with(france.boot.bbt, point - 1.96 * se)
france.boot.bbt$upper <- with(france.boot.bbt, point + 1.96 * se)

# Display the resulting data frame
france.boot.bbt
```

## 2.4 Ordered Optimal Classification
```{r message=FALSE, warning=FALSE}
library(asmcjr)
data(ANES2004_OOC)
issuescales <- ANES2004[,1:14]
head(issuescales)
```
The command below performs OOC on the 2004 ANES issue scale data in two dimensions:

```{r message=FALSE, warning=FALSE, include=FALSE}
# devtools::install_github('tzuliu/ooc')
library(ooc)
ooc.result <-ooc(issuescales,dims=2,minvotes=10,lop=0.001,
                 polarity=c(1,1),iter=25,nv.method="svm.reg",cost=1)
```

```{r}
issue.result <- ooc.result$issues.unique

# Set the row names of issue.result using the column names from issuescales
rownames(issue.result) <- colnames(issuescales)

# Print selected columns from issue.result
print(issue.result[, c("normVectorAngle2D", "wrongScale", 
                       "correctScale", "errorsNull", "PREScale")])

```


## 2.5 Using Anchoring Vignettes
calculate party means and standard deviations across all CHES experts
```{r}
# Load the asmcjr package
library(asmcjr)

# Load the ches_eu dataset
data(ches_eu)

# Calculate the column means of sub.europe, ignoring NA values
means <- colMeans(sub.europe, na.rm = TRUE)

# Calculate the standard deviations of the columns in sub.europe, ignoring NA values
sds2 <- apply(sub.europe, 2, sd, na.rm = TRUE)

# Convert sub.europe to a matrix
sub.europe <- as.matrix(sub.europe)

# Ensure the mode of sub.europe is numeric
mode(sub.europe) <- "numeric"

```

Call the blackbox_transpose 
```{r}
library(basicspace)
result <- blackbox_transpose(sub.europe,dims=3,
                             minscale=5,verbose=TRUE)
```

```{r}
# Create a data frame europe.dat containing x, y coordinates, means, party names, and types
europe.dat <- data.frame(
  x = -result$stimuli[[2]][,2],  # Negate the second column of stimuli and assign to x
  y = result$stimuli[[2]][,3],   # Use the third column of stimuli for y
  means = means,                 # Add the means calculated earlier
  party = colnames(sub.europe),  # Use the column names of sub.europe as party names
  type = car:::recode(means,     # Recode means into categories: Left, Moderate, and Right
    "lo:3 = 'Left'; 3:7 = 'Moderate'; 7:hi = 'Right'",
    as.factor = TRUE)
)

# Separate europe.dat into two data frames: parties.dat and vignette.dat
parties.dat <- europe.dat[-(1:3), ]  # Exclude the first three rows for parties.dat
vignette.dat <- europe.dat[(1:3), ]  # Include only the first three rows for vignette.dat

# Extract fit values from the result object
onedim <- result$fits[1, 3]  # Extract the one-dimensional fit value
twodim <- result$fits[2, 3]  # Extract the two-dimensional fit value
```



```{r message=FALSE, warning=FALSE, fig.cap="FIGURE 2.16: Result of Blackbox Transpose with Anchoring Vignettes"}
library(ggplot2)
ggplot(parties.dat, aes(x = x, y = y)) +
  geom_point(aes(shape = type, color = type), size = 3) +  
  
  scale_color_manual(values = gray.palette(3)) +  
  theme_bw() +  
  # Add text labels "A", "B", and "C" to the points in vignette.dat
  geom_text(data = vignette.dat, label = c("A", "B", "C"),
            show.legend = FALSE, size = 10, color = "black") +
  xlab(paste0("First Dimension (fit = ", round(onedim, 1), "%)")) +
  ylab(paste0("Second Dimension (fit = ", round(twodim, 1), "%)")) +
  theme(legend.position = "bottom", aspect.ratio = 1) +
  labs(colour = "Party Group", shape = "Party Group")

```



```{r message=FALSE, warning=FALSE, fig.cap="FIGURE 2.17: Result of Blackbox Transpose versus Mean Placementes"}
ggplot(parties.dat, aes(x=x, y=means)) +
  geom_smooth(method="loess", color="black", lwd=.5, se=FALSE) +
  geom_point(aes(shape=type, color=type), size=3) +
  scale_color_manual(values=gray.palette(3)) +
  theme_bw() +
  xlab("First Dimension Coordinates") +
  ylab("Mean Party Placement") +
  theme(legend.position="bottom", aspect.ratio=1) +
  labs(shape="Party Group", colour="Party Group")
```

