---
title: "Unfolding Analysis of Binary Choice Data"
# author: "Yen-Chieh Liao"
# date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Unfolding Analysis of Binary Choice Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## 5.3 Parametric Methods - NOMINATE
#### 5.3.3 Accessing DW-NOMINATE Scores


```{r}
library(asmcjr)
data(rcx)
```

```{r}
# select the Congress terms of interest
ncong <- c(90, 100, 110)

# Filter the data for rows where the Congress number is in ncong and dist is not equal to 0
sub <- with(rcx, which(cong %in% ncong & dist != 0))

# Create a matrix containing the first dimension scores (dwnom1), party affiliation, 
# and a factor indicating the Congress (e.g., "House 90", "House 100", "House 110")
polarization <- cbind(
  rcx[sub, c("dwnom1", "party")],  # Select columns for first dimension score and party
  congress = factor(paste("House", rcx$cong[sub]),  # Create a factor for the Congress
                    levels = c("House 90", "House 100", "House 110"))
)
```



```{r warning=TRUE, fig.cap="FIGURE 5.2: Partisan Polarization in the 90th, 100th, and 110th Houses"}
library(ggplot2)

# Ensure that 'party' is a factor
polarization$party <- as.factor(polarization$party)

# Create the plot
ggplot(polarization, aes(x = dwnom1, group = party, colour = party, fill = party)) +
  geom_density(adjust = 2.5, alpha = .2) +
  facet_wrap(~congress, ncol = 2) +
  xlab("DW-Nominate\n(First Dimension)") +
  theme_bw() +
  scale_colour_manual(values = c("gray25", "gray75"), name = "Party") +
  scale_fill_manual(values = c("gray25", "gray75"), name = "Party") +
  theme(aspect.ratio = 1, legend.position = "bottom")

```

#### 5.3.5 Example 1: The 108th US House
```{r message=FALSE, warning=FALSE}
library(wnominate)
result <- wnominate(hr108, ubeta=15, uweights=0.5, dims=2,
                    minvotes=20, lop=0.025, trials=1, polarity=c(1,5), verbose=FALSE)
```

Create a grouping variable based on partyCode and icpsrState.
```{r}
group <- rep(NA, nrow(result$legislators))
group[with(result$legislators, which(partyCode == 100 & icpsrState %in% c(40:51, 53, 54)))] <- 1
group[with(result$legislators, which(partyCode == 100 & icpsrState %in% c(1:39, 52, 55:82)))] <- 2
group[with(result$legislators, which(partyCode == 200))] <- 3
group[with(result$legislators, which(partyCode == 328))] <- 4

# Convert the group variable to a factor with appropriate labels
group <- factor(group, levels = 1:4,
                labels = c("Southern Dems", "Northern Dems",
                           "Republicans", "Independents"))

```

```{r warning=TRUE, fig.cap="FIGURE 5.3: W-NOMINATE Ideal Point Estimates of Members of the 108th US House of Representatives"}
library(ggplot2)
# Plot using plot_wnom_coords with customized colors and shapes
plot_wnom_coords(result, shapeVar = group, dropNV = FALSE) +
  scale_color_manual(values = gray.colors(4, end = .75), name = "Party Group",
                     labels = c("Southern Dems", "Northern Dems",
                                "Republicans", "Independents")) +
  scale_shape_manual(values = c("S", "N", "R", "I"), name = "Party Group",
                     labels = c("Southern Dems", "Northern Dems",
                                "Republicans", "Independents")) +
  theme_bw() +
  theme(aspect.ratio = 1, legend.position = "bottom") +
  xlab("First Dimension") +
  ylab("Second Dimension") +
  guides(colour = guide_legend(nrow = 2))
```

Shows the roll call choices of all voting legislators. 
```{r warning=TRUE, fig.cap="FIGURE 5.4: W-NOMINATE Analysis of the 108th House Vote on the Partial-Birth Abortion Ban Act of 2003, All Legislators"}
# Calculate the weight based on the result's weights
weight <-  result$weights[2] / result$weights[1]

# Create a data frame with the first and second dimension coordinates, adjusting the second dimension by the weight
wnom.dat <- data.frame(
  coord1D = result$legislators$coord1D,
  coord2D = result$legislators$coord2D * weight,
  group = group
)

# Plot the roll call results
plot_rollcall(result, hr108, wnom.dat, 528, wnom.dat$group, dropNV = TRUE) +
  theme_bw() +
  scale_shape_manual(values = c("S", "D", "R", "I"), name = "Party Group") +
  xlab("First Dimension") +
  ylab("Second Dimension") +
  theme(aspect.ratio = 1, legend.position = "bottom") +
  guides(shape = guide_legend(nrow = 2))

```

We can calculate the PRE (Proportional Reduction in Error) statistic and display it in the plot below the number of errors as below:
```{r}
rc.errors(result, hr108, 528)[c("tot.errors", "PRE")]
```


To plot only those legislators who committed voting errors, we can use the plot_rollcall() function from above.
```{r warning=TRUE, fig.cap="FIGURE 5.5: W-NOMINATE Analysis of the 108th House Vote on the Partial-Birth Abortion Ban Act of 2003, Errors Only"}
plot_rollcall(result, hr108, wnom.dat, 528,
             shapeVar=wnom.dat$group, onlyErrors=TRUE) +
  scale_shape_manual(values=c("S", "N", "R"), name="Party Group") +
  theme_bw() +
  theme(aspect.ratio=1, legend.position="bottom") +
  xlab("First Dimension") +
  ylab("Second Dimension") +
  guides(shape = guide_legend(nrow=2))
```
#### 5.3.5.3 The Coombs Mesh and Cutting Line Angles
The Coombs mesh also illustrates how cutting planes intersect to form polytopes (bounded regions in the space). 
```{r}
angles <- makeCutlineAngles(result)
head(angles)
```

```{r}
print(angles[528,])
```

```{r}
mean(abs(angles$angle), na.rm=TRUE)
```


```{r warning=TRUE, fig.cap="FIGURE 5.7: Histogram of Cutting Line Angles of Roll Call Votes in the 108th US House"}
ggplot(angles, aes(x=angle)) +
     geom_histogram() +
     theme_bw() +
     theme(aspect.ratio=1) +
     xlab("Cutline Angles")
```


#### 5.3.6 Example 2: The First European Parliament (Using the Parametric Bootstrap)
```{r message=FALSE, warning=FALSE}
library(wnominate)
result <- wnominate(rc_ep, ubeta=15, uweights=0.5, dims=2, minvotes=20,
                    lop=0.025, trials=5, polarity=c(25,25), verbose=FALSE)
```

```{r warning=TRUE, fig.cap="FIGURE 5.8: W-NOMINATE Ideal Point Estimates of Members of the First European Parliament"}
library(ggplot2)
plot_wnom_coords(result, shapeVar = result$legislators$MS, 
                 dropNV = FALSE, ptSize = 4, ci = FALSE) +
  scale_shape_discrete(name = "Party Group")
```


## 5.4 Nonparametric Methods - Optimal Classification
#### 5.4.2 Example 1: The French National Assembly during the Fourth Republic

```{r}
library(asmcjr)
data(france4)
rc <- rollcall(data=france4[,6:ncol(france4)],
   yea=1,
   nay=6,
   missing=7,
   notInLegis=c(8,9),
   legis.names=paste(france4$NAME,france4$CASEID,sep=""),
   vote.names=colnames(france4[6:ncol(france4)]),
   legis.data=france4[,2:5],
   vote.data=NULL,
   desc="National Assembly of the French Fourth Republic")
```

```{r message=FALSE, warning=FALSE}
library(oc)
result1 <- oc(rc, dims=1, minvotes=20, lop=0.025,
     polarity=2, verbose=FALSE)
result2 <- oc(rc, dims=2, minvotes=20, lop=0.025,
              polarity=c(2,2), verbose=FALSE)
```

<!-- ```{r} -->
<!-- fits <- cbind(result1$fits, result2$fits) -->
<!-- colnames(fits) <- c("1 Dim", "2 Dim") -->
<!-- rownames(fits) <- c("% Correct", "APRE") -->

<!-- # Convert degrees to radians -->
<!-- deg2rad <- function(x) x * pi / 180  -->
<!-- # 45 degrees in radians -->
<!-- rad45 <- deg2rad(45) -->
<!-- # Calculate 45-degree rotation matrix -->
<!-- A <- matrix(c(cos(rad45), -sin(rad45), -->
<!--               sin(rad45), cos(rad45)), -->
<!--             nrow = 2, ncol = 2, byrow = TRUE)   -->
<!-- ``` -->


<!-- ```{r warning=TRUE, fig.cap="FIGURE 5.10: Optimal Classification Ideal Point Estimates of Deputies of the French Fourth Republic"} -->
<!-- pb <- rc$legis.data$PAR -->
<!-- pb <- car::recode(pb, '1="Communitst"; 2="Socialists"; -->
<!--                   5="Christian Dems"; 7 = "Poujadists"; else=NA', -->
<!--       as.factor=TRUE) -->

<!-- plot_oc_coords(result, pb, dropNV=, ptSize=3, rotMat=A) + -->
<!--    theme_bw() + -->
<!--    coord_fixed() + -->
<!--    scale_shape_manual(values=c("C", "S", "D", "P"), -->
<!--        labels= c("Communist", "Socialist", -->
<!--                  "Christian Dem", "Poujadists"), -->
<!--        name = "Party") + -->
<!--    scale_color_manual(values=gray.colors(5), -->
<!--        labels= c("Communist", "Socialist", -->
<!--                  "Christian Dem", "Poujadists"), -->
<!--        name = "Party")+ -->
<!--  labs(x="First Dimension", y="Second Dimension") + -->
<!--  theme(aspect.ratio=1, legend.position="bottom") -->
<!-- ``` -->


